---
title: "Google Timeline Data Preview"
author: "Hong Gao"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M')` EST"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# © 2016 Hong Gao, NYU mHealth, All Rights Reserved

##################
# load libraries #
##################

if(!require("jsonlite")){
  install.packages("jsonlite")
  library("jsonlite")
} else {
  library("jsonlite")
}
if(!require("plyr")){
  install.packages("plyr")
  library("plyr")
} else {
  library("plyr")
}
if(!require("ggplot2")){
  install.packages("ggplot2")
  library("ggplot2")
} else {
  library("ggplot2")
}
if(!require("devtools")){
  install.packages("devtools")
  library("devtools")
} else {
  library("devtools")
}
if(!require("Mobility")){
  install_github("nyu-mhealth/Mobility")
  library("Mobility")
} else {
  library("Mobility")
}
if(!require("rgdal")){
  install.packages("rgdal")
  library("rgdal")
} else {
  library("rgdal")
}
if(!require("ggmap")){
  install.packages("ggmap")
  library("ggmap")
} else {
  library("ggmap")
}
if(!require("leaflet")){
  install.packages("leaflet")
  library("leaflet")
} else {
  library("leaflet")
}
if(!require("magrittr")){
  install.packages("magrittr")
  library("magrittr")
} else {
  library("magrittr")
}
if(!require("RColorBrewer")){
  install.packages("RColorBrewer")
  library("RColorBrewer")
} else {
  library("RColorBrewer")
}
```

### Google Timeline Data Source

Locations are recorded through Google Map or Google apps and shared with us securely. Downloaded data 
is in zip file, need to unzip and extract the json files. 

```{r message=FALSE, warning=FALSE}
# define file path
file.path <- "~/Downloads/test"
# change directory to file path
setwd(file.path)
# get all the file names in the folder
file.name <- list.files()
file.name <- file.name[file.name != "Takeout"]

# load all the files with a loop
gt.track <- NULL
for (i in file.name) {
  location <- jsonlite::fromJSON(unzip(i, 
                                     files = "Takeout/Location History/LocationHistory.json"))
  location <- location$locations
  # remove the .zip in file name and append it as a unique id
  location$uid <- substr(i, 1, nchar(i)-4)
  gt.track <- rbind(gt.track, location)
}

# change time and location format
gt.track$timestampMs <- as.numeric(gt.track$timestampMs)/1000
gt.track$datetime <- as.POSIXct(gt.track$timestampMs, origin="1970-01-01")
gt.track$latitude <- gt.track$latitudeE7/1E7
gt.track$longitude <- gt.track$longitudeE7/1E7

# give each person an unique id number
gt.track <- transform(gt.track, id = as.numeric(interaction(uid, drop=TRUE)))
gt.track <- gt.track[order(gt.track$id, gt.track$datetime),]
```

### Descriptives 

####Total number of points per person

```{r message=FALSE, warning=FALSE}
total.track <- count(gt.track, c("uid", "id"))
total.track
```

####Number of points per person by day

```{r message=FALSE, warning=FALSE}
# generate dates
gt.track$date <- as.Date(gt.track$datetime, format = "%Y-%m-%d")
day.track <- count(gt.track, c("uid", "id", "date"))
day.track
```

Plots of points per day

```{r message=FALSE, warning=FALSE}
# loop through each id
ids <- unique(gt.track$id)
for (i in ids) {
  p <- ggplot(gt.track[gt.track$id == i,], aes(date))
  p <- p + geom_bar() + 
    labs(title = paste0("id = ", i)) + 
    theme(panel.background=element_blank()) + 
    theme(panel.background= element_rect(color="black"))
  print(p)
}
```

####Calculating stay events

Stay events are defined as being within 50 meters of the previous GPS point for at least 10 minutes. Both 
distance and time thresholds are adjustable. 

```{r message=FALSE, warning=FALSE}
gt.track <- stayevent(gt.track, coor = c("longitude", "latitude"), 
                      time = "datetime", dist.threshold = 50, 
                      time.threshold = 10, time.units = "mins", 
                      groupvar = "id")
# aggragate a data frame with only stayevents
gt.stay <- unique(gt.track[c("id", "uid", "stayeventgroup", 
                             "date", "stayeventlon", "stayeventlat")])
gt.stay <- gt.stay[!is.na(gt.stay$stayeventlat),]
#gt.stay <- subset(gt.stay, stayeventlon < -70 & gt.stay$stayeventlat < 41)
```


####Visualizing stay events

**Raw tracks and stay events**

The map contains two layers showing raw mobility tracks and calculated stay events, which can 
be turned on and off through the layer control on the upper right corner. The color scheme 
indicates different individuals. 


```{r message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
# transform stay events from data frame to spatial points
coordinates(gt.stay) <- c("stayeventlon", "stayeventlat")
# extract bounding box of the points
gt.stay.box <- as.numeric(bbox(gt.stay))
names(gt.stay.box) <- c("left", "bottom", "right", "top")
# plot
# map<- ggmap(get_map(location = gt.stay.box, source = "stamen", color = "bw"))
# ggmap(get_stamenmap(gt.stay.box, zoom = 12, maptype = "toner-lite"))
gt.stay <- data.frame(gt.stay)
gt.stay$id_factor <- as.factor(gt.stay$id)
qmplot(stayeventlon, stayeventlat, data = gt.stay, maptype = "toner-lite", 
       color = gt.stay$id_factor, zoom = 12, legend = F)
```

```{r message=FALSE, warning=FALSE}
# use leaflet package to produce an interactive map
pal1 <- colorFactor("RdYlBu", domain = levels(gt.stay$uid))
pal2 <- colorFactor(c("firebrick2", "blue2"), domain = levels(gt.track$uid))
# leaflet(data = gt.stay) %>% addProviderTiles(providers$CartoDB.Positron) %>%
#   addCircleMarkers(~stayeventlon, ~stayeventlat, radius = 6, 
#              color = ~pal1(uid), stroke = FALSE, fillOpacity = 0.7, 
#              popup = ~as.character(uid), label = ~as.character(uid))

leaflet(data = gt.stay) %>% addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(data = gt.track, ~longitude, ~latitude, radius = 4, 
                   group = "raw points", 
             color = ~pal1(uid), stroke = FALSE, fillOpacity = 0.6, 
             popup = ~as.character(uid), label = ~as.character(datetime)) %>% 
    addCircleMarkers(~stayeventlon, ~stayeventlat, radius = 6, group = "stayevents", 
             color = ~pal2(uid), stroke = FALSE, fillOpacity = 0.7, 
             popup = ~as.character(uid), label = ~as.character(uid)) %>% 
  addLayersControl(
    overlayGroups = c("stayevents", "raw points"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


**Stay events by dates**

Visualize all the stay events for id = 2 by dates. Dates can be turned on and off as well. 


```{r message=FALSE, warning=FALSE}
# by dates
gt.stay1 <- gt.stay[gt.stay$id == 2,]
gt.stay1$date1 <- as.factor(gt.stay1$date)

quakes <- quakes %>%
  dplyr::mutate(mag.level = cut(mag,c(3,4,5,6),
                                labels = c('>3 & <=4', '>4 & <=5', '>5 & <=6')))

gt.stay1.df <- split(gt.stay1, gt.stay1$date1)

l <- leaflet() %>% addProviderTiles(providers$CartoDB.Positron)

names(gt.stay1.df) %>%
  purrr::walk( function(df) {
    l <<- l %>%
      addMarkers(data = gt.stay1.df[[df]],
                          lng=~stayeventlon, lat=~stayeventlat,
                          label=~as.character(uid),
                          popup=~as.character(date),
                          group = df,
                          clusterOptions = markerClusterOptions(removeOutsideVisibleBounds = F),
                          labelOptions = labelOptions(noHide = F,
                                                       direction = 'auto'))
  })

l %>%
  addLayersControl(
    overlayGroups = names(gt.stay1.df),
    options = layersControlOptions(collapsed = FALSE)
  )
```

####Calculate stay points

Stay points are defined as the latent locations people have been to. They are extracted 
based on stay events but without the temporal element. With stay points defined, we can 
decide if multiple stay events happen at the same place. Further, we can calculate return 
probabilities and rank order places people have been to. 

**test kmeans method**

```{r message=FALSE, warning=FALSE}
# in order to avoid longer processing time if stay events are scattered far away, 
# stay events that are clustered using kmeans method are processed together
lat.id1 <- gt.stay1$stayeventlat
lon.id1 <- gt.stay1$stayeventlon
# run kmeans for 2-15 centers and see which number results in the smallest withinss
cluster.summary <- NULL
set.seed(20)
for (i in 2:15){
  km <- kmeans(cbind(lat.id1, lon.id1), centers = i, nstart = 20)
  cluster.i <- c(i, sum(km$withinss))
  cluster.summary <- data.frame(rbind(cluster.summary, cluster.i))
}
cluster.n <- cluster.summary[cluster.summary$X2 == min(cluster.summary$X2),]$X1
# rerun kmeans clustering using cluster.n as the number of clusters
km <- kmeans(cbind(lat.id1, lon.id1), centers = cluster.n, nstart = 20)
# merge cluster number back to the points
gt.stay1$cluster.n <- km$cluster
# visualize the clusters
pal3 <- colorRampPalette(brewer.pal(8, "Accent"))
pal4 <- colorFactor(pal3(cluster.n), domain = levels(gt.stay1$cluster.n))
leaflet(data = gt.stay1) %>% addProviderTiles(providers$CartoDB.Positron) %>% 
    addCircleMarkers(~stayeventlon, ~stayeventlat, radius = 6, group = "stayevents", 
             color = ~pal4(cluster.n), stroke = FALSE, fillOpacity = 0.7, 
             popup = ~as.character(uid), label = ~as.character(date))
```

**Loop through clusters to find stay points**

```{r message=FALSE, warning=FALSE}

```

